Annotated example of index standardization with VAST

### 2021-12-15 doc/appendix-pcod-VAST.Rmd
Edit delta-gamma compare

# effort is 1 when using CPUE instead of observed weight as the response:
pcod$effort <- 1
pcod <- as.data.frame(pcod) # ensure not a tibble
f <- here("doc", "appendix-VAST", "vast-cache.rds")
if (!file.exists(f)) {
  fit <- fit_model(
    settings = settings,
    Lat_i = pcod[, "lat"],
    Lon_i = pcod[, "lon"],
    t_i = pcod[, "year"],
    b_i = pcod[, "density"],
    a_i = pcod[, "effort"],
    input_grid = input_grid,
    working_dir = paste0(here("doc", "appendix-VAST"), "/")
  )
  saveRDS(fit, file = f)
} else {
  fit <- readRDS(f)
}
```

@@

```{r sdmTMB-example, echo = TRUE, cache=TRUE}
mesh <- make_mesh(pcod, xy_cols = c("X", "Y"),
  mesh = fit$spatial_list$MeshList$isotropic_mesh)
m1 <- sdmTMB(present ~ 0 + as.factor(year), 
  data = pcod, mesh = mesh, family = binomial(link = "logit"),
  time = "year", spatiotemporal = "iid", spatial = "on")
m1

dat2 <- subset(pcod, density > 0)
mesh2 <- make_mesh(dat2, xy_cols = c("X", "Y"),
  mesh = fit$spatial_list$MeshList$isotropic_mesh)
m2 <- sdmTMB(density ~ 0 + as.factor(year), 
  data = dat2, mesh = mesh2, family = Gamma(link = "log"),
  time = "year", spatiotemporal = "iid", spatial = "on")
m2
```

@@

```{r sdmTMB-sim-index, cache=TRUE, echo=TRUE}
set.seed(28239)
p_bin <- predict(m1, newdata = qcs_grid, sims = 1000)
p_pos <- predict(m2, newdata = qcs_grid, sims = 1000)

p_bin_prob <- plogis(p_bin)
p_pos_exp <- exp(p_pos)
p_combined <- log(p_bin_prob * p_pos_exp)
ind_combined_sims <- get_index_sims(p_combined, area = rep(4, nrow(p_combined)))
```

Now, we can compare the indices. We will read in the VAST index from the `Index.csv` file created by the `plot.fit_model()` method above.

```{r index-compare, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
vast_i <- read.csv(here("doc/appendix-VAST/Index.csv")) %>%
  mutate(index = "VAST", year = as.numeric(Time), est = Estimate, 
    se = Std..Error.for.ln.Estimate.) %>% 
  select(index, year, est, se) %>% 
  mutate(lwr = exp(log(est) + qnorm(0.025) * se)) %>% 
  mutate(upr = exp(log(est) + qnorm(0.975) * se))
sdm_i <- ind_combined_sims %>% mutate(index = "sdmTMB")
both_i <- bind_rows(sdm_i, vast_i) %>% filter(est > 0)
ggplot(both_i, aes(x = year, y = est, ymin = lwr, ymax = upr, colour = index)) + 
  geom_ribbon(alpha = 0.3) +
  geom_line() + 
  ylim(0, max(both_i$upr)) +
  coord_cartesian(expand = FALSE)
```

>>>>

# effort is 1 when using CPUE instead of observed weight as the response:
pcod$effort <- 1
pcod <- as.data.frame(pcod) # ensure not a tibble
f <- here("doc", "appendix-VAST", "vast-cache.rds")
if (!file.exists(f)) {
  tictoc::tic()
  fit <- fit_model(
    settings = settings,
    Lat_i = pcod[, "lat"],
    Lon_i = pcod[, "lon"],
    t_i = pcod[, "year"],
    b_i = pcod[, "density"],
    a_i = pcod[, "effort"],
    input_grid = input_grid,
    working_dir = paste0(here("doc", "appendix-VAST"), "/")
  )
  tictoc::toc()
  saveRDS(fit, file = f)
} else {
  fit <- readRDS(f)
}
```

@@

```{r sdmTMB-example, echo = TRUE, cache=TRUE}
mesh <- make_mesh(pcod, xy_cols = c("X", "Y"),
  mesh = fit$spatial_list$MeshList$isotropic_mesh)
tictoc::tic()
m1 <- sdmTMB(present ~ 0 + as.factor(year), 
  data = pcod, mesh = mesh, family = binomial(link = "logit"),
  time = "year", spatiotemporal = "iid", spatial = "on")
tictoc::toc()
m1

dat2 <- subset(pcod, density > 0)
mesh2 <- make_mesh(dat2, xy_cols = c("X", "Y"),
  mesh = fit$spatial_list$MeshList$isotropic_mesh)
tictoc::tic()
m2 <- sdmTMB(density ~ 0 + as.factor(year), 
  data = dat2, mesh = mesh2, family = Gamma(link = "log"),
  time = "year", spatiotemporal = "iid", spatial = "on")
tictoc::toc()
m2
```

@@

```{r sdmTMB-sim-index, cache=TRUE, echo=TRUE}
tictoc::tic()
set.seed(28239)
p_bin <- predict(m1, newdata = qcs_grid, sims = 500)
p_pos <- predict(m2, newdata = qcs_grid, sims = 500)

p_bin_prob <- plogis(p_bin)
p_pos_exp <- exp(p_pos)
p_combined <- log(p_bin_prob * p_pos_exp)
ind_combined_sims <- get_index_sims(p_combined, area = rep(4, nrow(p_combined)))
tictoc::toc()
```

Now, we can compare the indices. 
<!-- We will read in the VAST index from the `Index.csv` file created by the `plot.fit_model()` method above. -->

```{r index-compare, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
vast_i1 <- read.csv(here("doc/appendix-VAST/Index.csv")) %>%
  mutate(index = "VAST", year = as.numeric(Time), est = Estimate, 
    se = Std..Error.for.ln.Estimate.) %>% 
  select(index, year, est, se) %>% 
  mutate(lwr = exp(log(est) + qnorm(0.025) * se)) %>% 
  mutate(upr = exp(log(est) + qnorm(0.975) * se))
sdm_i <- ind_combined_sims %>% mutate(index = "sdmTMB")
both_i <- bind_rows(sdm_i, vast_i) %>% filter(est > 0)
ggplot(both_i, aes(x = year, y = est, ymin = lwr, ymax = upr, colour = index)) + 
  geom_ribbon(alpha = 0.3) +
  geom_line() + 
  ylim(0, max(both_i$upr)) +
  coord_cartesian(expand = FALSE)
```

```{r}
plot(ind_sdmTMB$est, vast_i$est);abline(0, 1)
plot(ind_sdmTMB$upr, vast_i$upr);abline(0, 1)
plot(ind_sdmTMB$lwr, vast_i$lwr);abline(0, 1)

(ind_sdmTMB$est - vast_i$est) / vast_i$est
(ind_sdmTMB$upr - vast_i$upr) / vast_i$upr
(ind_sdmTMB$lwr - vast_i$lwr) / vast_i$lwr
```
